Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-11-15 grove-paper.tex
make code in prose match the diagrams
also fix a few small grammar errors and add some whitespace to latex
source

% $\arraycolsep=2pt\begin{array}{llcl}
% \mathsf{Typ} & \tau & ::= & t ~\vert~ \aparr{\tau}{\tau} ~\vert~ \aall{t}{\tau} ~\vert~ \arec{t}{\tau} ~\vert~ \aprod{\labelset}{\mapschema{\tau}{i}{\labelset}} ~\vert~ \asum{\labelset}{\mapschema{\tau}{i}{\labelset}}\\
% \mathsf{Exp} & e & ::= & x ~\vert~ \aelam{\tau}{x}{e} ~\vert~ \aeap{e}{e} ~\vert~ \aetlam{t}{e} ~\vert~ \aetap{e}{\tau} ~\vert~ \aefold{e} ~\vert~ \aetpl{\labelset}{\mapschema{e}{i}{\labelset}} ~\vert~  \aein{\ell}{e} \\
% & & \vert & \aematchwith{n}{e}{\seqschemaX{r}}\\
% \mathsf{Rule} & r & ::= & \aematchrule{p}{e}\\
% \mathsf{Pat} & p & ::= & x  ~\vert~ \aewildp ~\vert~ \aefoldp{p} ~\vert~ \aetplp{\labelset}{\mapschema{p}{i}{\labelset}} ~\vert~ \aeinjp{\ell}{p}
% \end{array}$

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \_                                                            \\
      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \_ \\
      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \_                               \\
    \end{array}
  \]
  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
  \label{fig:Syntax as a grammar}
\end{figure}

@@

The language we are modeling is shown in \autoref{fig:Syntax as a grammar}.
Standard for a simply-typed lambda calculus except that
all terms have holes (represented by $\_$)
and the argument to a lambda is its own term.
The latter of which allows us to express lambdas with a hole for the bound variable.

In order to model this we define

Turn that into a constructor set K + index set I + arity function (has type: K -> $\wp$(I)).

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{ll}
      \arity : \K \rightarrow & \wp(\I)                                        \\
      \hline
      \arity(\Root)={}        & \left\{ \Root \right\}                         \\
      \arity(\PatVar)={}      & \left\{ \right\}                               \\
      \arity(\ExpVar)={}      & \left\{ \right\}                               \\
      \arity(\ExpLam)={}      & \left\{ \LamParam, \LamType, \LamBody \right\} \\
      \arity(\ExpApp)={}      & \left\{ \AppFun, \AppArg \right\}              \\
      \arity(\ExpPlus)={}     & \left\{ \PlusLeft, \PlusRight \right\}         \\
      \arity(\TypArrow)={}    & \left\{ \ArrowArg, \ArrowResult \right\}       \\
      \arity(\TypNum)={}      & \left\{ \right\}                               \\
    \end{array}
  \]
  \caption{Constructors, Indexes and Arity}
  \label{fig:Constructors, Indexes and Arity}
\end{figure}

@@

TODO: rename indices to positions

TODO: look into using bold plus and minus (google results)

@@

A vertex $v = (u, k)$ identifies a distinct instance of the constructor $k$.

A constructor $k \in \K$ represents a different type of node in the target language's abstract syntax tree.

A position $p \in \P$ identifies a location, adjacent to some parent vertex.

Every graph contains a distinguished root vertex, root constructor, and root position, respectively named $v_0$, $k_0$, and $p_0$.

% Define $\sqcup_{edge state}$ (i.e. join)

% lattice ordering $\bot < + < -$ LUB e.g., join A B = C

You can map from syntax to graph by selecting
unique IDs $u$ (for both edges and vertexes).
- holes are not explicit in the graph

@@

We define the join operation $s_1 \sqcup s\_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering.
For example, $\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
and $\Plus \sqcup \Minus = \Minus$, which means that once an edge is deleted it can never be restored.
%smallest element of $\Sigma$ that is greater than or equal to both $s_1$ and $s_2$.

>>>>

% XXX "p" is overloaded - could mean pattern or position

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Patterns:}          & p    & {}\in Pat & {}::={} & x \mid \hole                                                            \\
      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x \mid \lambda p\mathord{:}\tau.e \mid e~e \mid e + e \mid n \mid \hole \\
      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow \tau \mid Num \mid \hole                               \\
    \end{array}
  \]
  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
  \label{fig:Syntax as a grammar}
\end{figure}

@@

The language we are modeling is shown in \autoref{fig:Syntax as a grammar},
standard for a simply-typed lambda calculus except that all types of terms have
holes (represented by $\hole$). In particular, the argument to a lambda can be
either a bound variable or a hole.

% In order to model this we define

% Turn that into a constructor set K + position set P + arity function (has type: K -> $\wp$(P)).

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{ll}
      \arity : \K \rightarrow & \wp(\P)                                        \\
      \hline
      \arity(\Root)={}        & \left\{ \Root \right\}                         \\
      \arity(\PatVar)={}      & \left\{ \right\}                               \\
      \arity(\ExpVar)={}      & \left\{ \right\}                               \\
      \arity(\ExpLam)={}      & \left\{ \LamParam, \LamType, \LamBody \right\} \\
      \arity(\ExpApp)={}      & \left\{ \AppFun, \AppArg \right\}              \\
      \arity(\ExpPlus)={}     & \left\{ \PlusLeft, \PlusRight \right\}         \\
      \arity(\TypArrow)={}    & \left\{ \ArrowArg, \ArrowResult \right\}       \\
      \arity(\TypNum)={}      & \left\{ \right\}                               \\
    \end{array}
  \]
  \caption{Constructors, Indexes and Arity}
  \label{fig:Constructors, Indexes and Arity}
\end{figure}

@@

@@

A vertex $v = (u, k)$ identifies a distinct instance of some constructor $k$,
where $k \in \K$ corresponds to a type of expression. Each contructor has an
arity, which we define as a set of positions. A position $p \in \P$ identifies a
location, adjacent to some parent vertex, from which edges to child vertices may
originate. The $\arity$ function in \autoref{fig:Constructors, Indexes and
  Arity} gives the complete mapping from expression constructors to potential edge
positions. Every graph contains a distinguished root vertex, root constructor,
and root position, respectively named $v_0$, $k_0$, and $p_0$.

A graph can be derived from an expression in the calculus by starting with an
empty graph, selecting unique ids $u, u_k$ and a constructor $k$ for the
outermost expression, and creating the edge $\e = (u, v_0, p_0, (u_k, k))$.
Then, for each sub-expression at some position $p \in \arity(k)$, select unique
id $u'$ and constructor $k'$, and create the edge $e' = (u', v, p, v')$, where
$v = (u_k, k)$ and $v' = (u', k')$. Continuing in this manner, treat each
sub-expression, starting with the ones at positions in $\arity(k')$, as if it
were the outermost and create an edge to the vertices representing each of its
immediate sub-expressions. Note that holes are not modeled explicitly in the
graph. When an expression has a hole, it means that the graph has no edges
originating from that position.

% You can map from syntax to graph by selecting
% unique IDs $u$ (for both edges and vertexes).
% - holes are not explicit in the graph

@@

We define the join operation $s_1 \sqcup s\_2$ to be the least upper bound of $s_1$ and $s_2$ with respect to the $\sqsubset$ ordering.
For example,
$\Plus \sqcup \bot = \Plus$, which means edges can be created if they have never existed,
and $\Plus \sqcup \Minus = \Minus = \Minus \sqcup \Minus$, which means that once an edge is deleted it can never be restored.

%smallest element of $\Sigma$ that is greater than or equal to both $s_1$ and $s_2$.
